Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory

### 2020-10-29 papers/thesis-kirstie/thesis.tex
edit-intro

The goal of this project is to generate temperature-density and 
pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using a computer program. 
To do this we need to find the Helmholtz free energies that the fluid
would have in a liquid state, and in a crystalline solid state 
at a given temperature and density. 
%Once the Helmholtz free energy for both the crystal and liquid states of 
%the fluid are found for a given temperature and density, they can be 
%compared to find the resulting state of the fluid. 
Outside the region of 
coexistence which exists between the liquid and crystal states, 
the fluid will be completely in a crystalline state if the 
crystal free energy is lower than the liquid free energy, and it will be 
completely in a liquid state if the liquid free energy is lower than 
the crystal free energy. Maxwell's Double Tangent Construction is used 
to determine the actual number densities of the crystal and liquid states
at the boundaries of the coexistance region for a given temperature from 
which the phase diagrams can be constructed.

In general, the Helmholtz free energy can be written as a functional of 
the number density profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
Classical Density Functional Theory (cDFT) is used to find the equilbrium 
Helmholtz free energy of the system 
by varying the number density profile $n(\vec{r})$ until the free energy 
is minimized. The minimum free energy, and its corresponding number 
density profile are then the equilibrium values of the Helmholtz free energy
of the system at the given temperature and number density.

To find the Helmholtz free energy both the ideal Helmholtz free energy
and the excess Helmholtz free energy must be found and added together.
The ideal Helmholtz free energy is the free energy the fluid would have in
the absence of any interactomic potential energy $V(r)$ between the spheres, 
or atoms that make up the fluid. The excess Helmholtz free energy is that 
additional energy which results from the presence of an interatomic potential 
energy $V(r)$ between the spheres.

@@

Soft Fundamental Measure Theory (SFMT) is a method used to form an expression 
for the excess Helmholtz free energy of a particular model fluid without 
having to take explicitly into account the interaction each sphere, or atom, 
has with every other sphere. 

@@

%which is known as the Weeks-Chandler-Anderson model fluid. 

@@

Schmidt applied his Soft Fundamental Measure Theory to various fluids 
with soft repulsive and soft attractive potentials, and developed an erf 
model for spheres with steep, short-range soft repulsive 
potentials~\cite{schmidt2000fluid}.

@@

In this paper, classical Density Functional 
Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
dependent parameters to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
Results will be compared to those generated from Monte-Carlo simulations.

>>>>

The goal of this project is to generate temperature-density and 
pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
model fluid using a computer program. 
To do this we need to find the Helmholtz free energies that the fluid
would have in a liquid state, and in a crystalline solid state 
at a given temperature and density. 
%Once the Helmholtz free energy for both the crystal and liquid states of 
%the fluid are found for a given temperature and density, they can be 
%compared to find the resulting state of the fluid. 
%
%save:
%Outside the region of 
%coexistence which exists between the liquid and crystal states, and which 
%consists of various amounts of liquid and crystal existing together,
%the fluid will be completely in a crystalline state if the 
%crystal free energy is lower than the liquid free energy, and it will be 
%completely in a liquid state if the liquid free energy is lower than 
%the crystal free energy. Maxwell's Double Tangent Construction is used 
%to determine the actual number densities of the crystal and liquid states
%at the boundaries of the coexistance region for a given temperature from 
%which the phase diagrams can be constructed.
%end save
%
%In general, if at some number density and temperature the crystal free energy 
%is lower than the homogeneous free energy of the liquid, then the fluid will 
%be in the crystalline state. If the homogeneous liquid free energy is lower 
%than the crystal free energy, then the fluid will be in the liquid state. 
%
%To find the equilibrium Helmholtz free energies we use classical Density 
%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
%%the corresponding equilibrium number density profile. 
%In particular, 
In general, the Helmholtz free energy, 
which consists of an ideal part and an excess part ,
can be written as a functional of 
the number density profile $n(\vec{r})$ which describes the way the average number 
density varies spatially. 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
%

In this paper, classical Density Functional Theory (cDFT)
will be used to find the equilibrium Helmholtz free energy, and 
Soft Fundamental Measure Theory (SFMT) will be used to form the functional for the excess Helmholtz 
free energy. 
%to predict the freezing of a Weeks-Chandler-Anderson 
%model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
%dependent parameters to form the functional for the excess Helmholtz free energy. 
%for generating the weight funcions, and Tarazona's tensor functional. 
%Classical Density Functional Theory (cDFT) is used to find the equilbrium 
According to classical Density Functional Theory (cDFT) the equilbrium 
Helmholtz free energy of the system 
can be found
by varying the number density profile $n(\vec{r})$ until the free energy 
is minimized. 
%The minimum free energy, and its corresponding number 
%density profile are then the equilibrium values of the Helmholtz free energy
%of the system at the given temperature and number density. 
%To find the Helmholtz free energy both the ideal Helmholtz free energy
%and the excess Helmholtz free energy must be found and added together.
%The ideal Helmholtz free energy is the free energy the fluid would have in
%the absence of any interactomic potential energy $V(r)$ between the spheres, 
%or atoms that make up the fluid. 
%The excess Helmholtz free energy is that 
%additional energy beyond the ideal Helmholtz free energy which results from the presence of an interatomic potential 
%energy $V(r)$ between the spheres.

@@

Soft Fundamental Measure Theory (SFMT) makes the task of forming an expression
%is a method used to form an expression 
for the excess Helmholtz free energy of a particular model fluid 
much easier since the interaction that each sphere, or atom, 
has with every other sphere does not have to be taken explicitly into account.
%without 
%having to take explicitly into account the interaction each sphere, or atom, 
%has with every other sphere which would make the task very difficult, 
%or even prohibtive.

@@

%which is known as the Weeks-Chandler-Anderson model fluid. f

@@

%Schmidt applied his Soft Fundamental Measure Theory to various fluids 
%with soft repulsive and soft attractive potentials, and developed an erf 
%model for spheres with steep, short-range soft repulsive 
%potentials~\cite{schmidt2000fluid}.

@@

Results will be compared to those from Monte-Carlo simulations.
